---
title: "<center>Homework 3</center>"
author: "<center>Input and Output</center>"
output: pdf_document
---

You are required to write a *markdown document* containing text, code, and at least one example. Thus you must create an Rmd file, that RStudio can convert to html. 

When writing `R` functions for this set of exercises, you may utilize any `R`
functions of your choice, except ones that perform identical or nearly
identical tasks relative to the functions detailed in the exercises.

1. Write an `R` function that accepts three arguments: characters input,
output, and an integer $n$, and then performs the following.  The function
reads the text file input, for example using the function `scan`,
identifies which words the text file contains, counts the number of times
each word occurs in the text, sorts the words according to their counts,
and puts the sorted words and their corresponding counts into a data frame
which is then written to the file `output`. The integer $n$ is an optional
argument that limits the output to the $n$ most common words.

    *Remark*: For testing purposes, you may use the file
    `HW3Supplement_const.txt` available on the CCLE webpage, which
    is a text file containing The Constitution of the United States.

2. The text file `usaircarrierRPM0374-0714.txt`, available on the CCLE webpage, contains
the monthly Revenue Passenger Miles (RPM), as reported by the U.S. air carriers to the Bureau of
Transportation Statistics, between March 1974 and July 2014.
Because the monthly RPM numbers exhibit seasonal fluctuations that easily could veil potential
changes in the underlying demand, and obscure the general market development, it is common to
smoothen such monthly fluctuations by applying a 12-month moving average filter. For each month,
the 12-month moving average filter is simply the average of the 12 months including and preceding
that month.
    (a) Write `R` code that reads the text file containing the RPM series, and produces the corresponding series of the 12-month moving averages of the RPM series.
    (b) Write code that, based on the moving average series, computes the average yearly growth during, respectively, the 70’s, 80’s, 90’s, 00’s and 10’s.  

    *hint*: Using the class `Date` may be practical.

3. Using the data of Exercise 2, write `R` code that produces a plot of the moving average series
overlaid on top of the original RPM series. Ideally, the plot should have a title, axis labels, and
a legend. For an accurate representation, make sure that the "y-axis" is scaled so that the number
zero is included. For reference, you make look at the file `HW3Supplement_RPMplot.pdf` on
the CCLE webpage.

4. Consider the following. If we in a new session write in the `R` console:
    ```{r}
    my_fun <- function(){
      my_object <- 0
      print(find("my_object"))
    }
```
and call `my_fun()`, then `R` will print the  
```
    character()
```
If we on the other hand in a new session write in the `R` console, and note the subtle difference:  
    ```{r}
    my_fun <- function(){
      my_object <<- 0
      print(find("my_object"))
    }
```
and call `my_fun()`, then `R` will print the following.  
```
[1] ".GlobalEnv"
```
Complete the following Parts:  
        (a) Explain, as precisely as you can, why the above described phenomenon occurs, and what the respective print-outs mean.  
        (b) We have mentioned that the line of code `my_object <- 0` is a short form of the function call `assign("my_object",0)`. The line of code `my_object <<- 0`, written above, is also short for a call of the `assign` function, but which call?  
        (c) While it is in general strongly discouraged, it is possible to reference an object created in an `R` session from within the body of a function, without passing the object as an argument. Explain how this can be done.  

Submit
------------------------------------------------
Upload your homework before the end of the due date via the *Homework* page
on the class CCLE. Submit only the Rmd file, but make sure RStudio knits it
correctly to html.
The `R` code
must be clearly readable and properly commented. Explain your solutions
well.



Cooperation
------------------------------------------------
A limited amount of cooperation on the homework is allowed, but everybody
hands in their own document. If you have cooperated in a group, indicate
this in your document and mention the names of the other students in the
group. Groups of more than four students are not allowed. Everybody in a
group will get the same grade, even if the individual documents are
different.
