---
title: "Homework1"
author: "Rmarkdown and Elementary programming"
date: "October 2, 2015"
output: html_document
---

Homework
Huang Mengyu
ID 204-707-588


1. Using only simple logical expressions, write an `R` function, for
example named `my_floor`, that accepts a floating-point number and returns
the greatest integer number that is less than or equal to the argument
floating-point number.

```{r}
my_floor<-function(a){
  i=0
  if(a>=i){
    while(a>=i){
      i=i+1
    }
    b=i-1
  }else{
    while(a<i){
      i=i-1
    }
    b=i
  }
  return(b)
}
my_floor(3.333)
my_floor(-0.175)
my_floor(99.99)
```


2. Using only simple logical expressions and/or your floor
function, write an `R` function, for example named `my_ceil`, that accepts a
floating-point number and returns the smallest integer number that is
greater than or equal to the argument floating-point number.

```{r}
my_ceil<-function(x){
  b<<-my_floor(x)
  if(b!=x){
    y=b+1
  }else{
    y=b
  }
  return(y)
}
my_ceil(4.75)
my_ceil(-1.472)
my_ceil(-99.999)
```


3. Using only simple logical expressions and/or your `R` functions,
write a function that accepts a floating-point number $x$ and an
integer-valued parameter $k$ that returns the floating-point number rounded
to $k$ decimals. A negative value of $k$ corresponds to rounding to a number
with $k$ trailing zeroes.

I don't really understand how to deal with the situation when k<0, so my function mainly deal with the situation that if k>=0

```{r}
my_round<-function(x,k){
  i=my_floor(x)
  while(abs(i-x)>=10^(-k)){
      i=i+10^(-k)
  }
  if(abs(i-x)>=abs(i+10^(-k)-x)){
    i=i+10^(-k)
   }
  return(i)
}
my_round(35.879092,4)
my_round(-33.425,1)
```

4. Using only simple logical expressions and/or your `R` functions,
write a function that accepts a non-negative integer number $x$ and a
positive integer $k$ that returns $x$ modulo $k$.

```{r}
modulo<-function(x,k){
  while(x>=k){
    x=x-k
  }
  return(x)
}
modulo(0,5)
modulo(3,5)
modulo(5,3)
modulo(16,3)
```


5. *Bubble sort* is a simple algorithm that sorts a vector or
numbers by evaluating consecutive pairs of numbers and swapping them if
the former number in the pair is greater than the latter. Implementing this
algorithm, write an `R` function, for example named `my_bubble` that accepts a
vector of floating-point numbers and returns the vector sorted from the
smallest number to the largest.

```{r}
bubble_sort<-function(x){
  for(i in 1:length(x)){
    for(j in i:length(x)){
      if(x[i]>x[j]) {
        a<-x[i]
        x[i]<-x[j]
        x[j]<-a
      }
    }
  }
  return(x)
}
x<-as.integer(runif(10)*100)
bubble_sort(x)
```


6. Suppose $x$ and $y$ are two non-zero double-precision floating
point numbers, one very large and the other very small. Explain why `R` under
some circumstances can return (erroneously) that the identity $x + y = x$ is
true.

All the real numbers in R are stored as double-precision. Though double-precision is 64-bit and it do 
provide more accuracy, it is still not possible for R to store all the decimals of a number because the
computer does not have an infinite storage. Thus, the double-precision numbers may still be just the 
approximation of numbers such as pi. When x is a very large number while y is rather small, the error may occur because of the approximation.