---
title: "Homework2"
author: "Mengyu Huang"
date: "October 26, 2015"
output: html_document
---
1. In statistics, samples collected are often affected by
*missing values*. For example, in a clinical trial it often
happens that patients do not complete the trial because they may perhaps
have to relocate to a different state, or similar, yielding a missing
value in the data set. Still, for scientific and economic reasons the
sample collected should in general be utilized even though one or a few
observations are incomplete.  

    There are several ways of addressing the problem of missing values. One
is to delete the observation from the sample, and another is
*imputation* of missing values. There are a number of
common techniques for imputation of missing values, including imputation
with a fixed value, such as zero, or imputation with the mean or median
value. In `R`, missing values are often encoded by
`NA`. *Hint*: Now is a good time to look up `NA` in `R`.
    (a) Write a function that accepts a numeric array or data frame, and
        returns the array or data frame with the observations (rows) that
        have missing values, encoded by `NA`, deleted.  

```{r}
#1(a)
delete_NA<-function(x){
  del_num<-NULL
  for(i in 1:dim(x)[1]){
    if(complete.cases(x[i,])==FALSE){
      del_num<-c(del_num,i)
    }
  }
  x<-x[-del_num,]
  return(x)
}

#generate test data
library("datasets")
data <- randu[1:40,]; data[(abs(data -0.4) < 0.05)] <- NA

#show how function works
delete_NA(data)
```

(b) Augment the function, by adding an argument, so that the function,
        instead of deleting observations (rows) with missing values, can
        perform imputation of the missing values with a fixed number, which
        is specified through an additional argument with default value equal
        to zero.

```{r}
#1(b)
aug_NA<-function(x,augment=0){
  newvec<-rep(augment,dim(x)[2])
  for(i in 1:dim(x)[1]){
    if(complete.cases(x[i,])==FALSE){
      x[i,]<-newvec
    }
  }
  return(x)
}

#generate test data
library("datasets")
data <- randu[1:40,]; data[(abs(data -0.4) < 0.05)] <- NA

#show how function works
aug_NA(data)
aug_NA(data,1)
```

(c) Augment the function with another option: imputation of the missing
        value with either the mean, median, or statistical mode of the
        variable (column) of the missing value. 
```{r}
#1(c)
#I write a function that enables you to choose to substitute the missing values with mean or median value.
stat_aug_NA<-function(x,mod='mean'){
  for(i in 1:dim(x)[1]){
    if(complete.cases(x[i,])==FALSE){
      for(j in 1:dim(x)[2]){
        if(is.na(x[i,j]==TRUE)){
          if(mod=='mean'){
            newval<-mean(x[,j],na.rm = TRUE)
          }else if(mod=='median'){
            newval<-median(x[,j],na.rm = TRUE)
          }
          x[i,j]<-newval
        }
      }
    }
  }
  return(x)
}

#generate test data
library("datasets")
data <- randu[1:40,]; data[(abs(data -0.4) < 0.05)] <- NA

#show how function works
stat_aug_NA(data)
```

2.This exercise is a continuation of Exercise 1. Augment your
imputation function by adding the option of imputation of the missing
value by the *nearest neighbor value*. The nearest neighbor
value is the corresponding value of the observation that is most similar
to the observation with the missing value. Evaluate similarity by the
Euclidean distance, which is defined for two vectors
$\vec{x}=(x_1,\dots,x_p)$ and $\vec{y} = (y_1,\dots,y_p)$ by
$$d(\vec{x},\vec{y}) = \sqrt{\sum_{k=1}^p(x_k-y_k)^2}.$$
That is, the most similar observation to $\vec{x}$ is the
observation that has the smallest Euclidean distance relative to
$\vec{x}$; its nearest neighbor.  

```{r}
#function code added the option of imputation by the *nearest neighbor value*.
nearest_aug_NA<-function(x,mod='mean'){
  for(i in 1:dim(x)[1]){
    if(complete.cases(x[i,])==FALSE){
      #to judge if the mode equals to 'nearst neighbor value'. If not, then the function is just the same as 1(c)
      if(mod!='nearest neighbor value'){
        for(j in 1:dim(x)[2]){
          if(is.na(x[i,j]==TRUE)){
            if(mod=='mean'){
              newval<-mean(x[,j],na.rm = TRUE)
            }else if(mod=='median'){
              newval<-median(x[,j],na.rm = TRUE)
            }
            x[i,j]<-newval
          }
        }
      }else{
        #If we choose to substitute the missing value with the *nearest neighbor value*
        distance=NULL
        #try to use the formula to calculate the distance
        for(k in 1:dim(x)[1]){
          sum=0
          for(j in 1:dim(x)[2]){
            if(!is.na(x[i,j])){
              sum<-sum+(x[k,j]-x[i,j])^2
            }
          }
          distance<-c(distance,sqrt(sum))
        }
        #delete the distance if this row has missing value, because it's meaningless to include the rows that have missing values.
        for(l in 1:dim(x)[1]){
          if(!complete.cases(x[l,])){
            distance[l]<-NA
          }
        }
        #try to find out which row has the smallest Euclidean distance relative to the seleted row
        n<-which.min(distance)
        x[i,]<-x[n,]
       }
    }
  }
  return(x)
}

#generate test data
library("datasets")
data <- randu[1:40,]; data[(abs(data -0.4) < 0.05)] <- NA

#try the new mode(imputation of missing value by the *nearest neighbor value*)
nearest_aug_NA(data,'nearest neighbor value')
```

3.Write a function that accepts an object of type `ordered factor`, and returns its median
level. Further, augment the function by adding the option of returning
the statistical mode of the factor. 
```{r}
#basic function
median_level<-function(x){
  if(length(levels(x))%%2){
    n<-levels(x)[ceiling(length(levels(x))/2)]
  }else{
    n<-c(levels(x)[length(levels(x))/2],levels(x)[length(levels(x))/2+1])
  }
  return(n)
}
#for those ordered factors whose level length are even numbers, we return both two levels in the medium.

#test the function
mons=c("Mar","Apr","Jan","Nov","Jan","Sep","Oct","Sep","Nov","Aug","Jan","Nov","Nov","Feb","May","Aug","Jul","Dec","Aug","Aug","Sep","Nov","Feb","Apr","Jun")
mons=factor(mons,levels=c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec",ordered=TRUE))
mons=ordered(mons)
median_level(mons)

fert = c(10,20,20,50,10,20,10,50,20)
fert = factor(fert,levels=c(10,20,50),ordered=TRUE)
median_level(fert)

#further augment of the basic function
stat_factor<-function(x,mod){
  if(mod=='median'){
    if(length(levels(x))%%2){
      n<-levels(x)[ceiling(length(levels(x))/2)]
    }else{
      n<-c(levels(x)[length(levels(x))/2],levels(x)[length(levels(x))/2+1])
    }
  }else if(mod=='mode'){
    n<-which.max(table(x))
  }
  return(n)
}

#test the function(try to get the statistical mode)
fert
stat_factor(fert,'mode')
```

4.
    (a) Suppose that the variable `w` has been assigned to
    represent an `R list`. Explain what the difference is
    between using (single) brackets, `w[1]`, and using double
    brackets, `w[[1]]`.  
    
A: For lists, we always use [[]] to select a single element we want. we can reference a list member directly. And if we use [], it returns a list of the selected elements, a list slice.

The [[]] form allows only a single element to be selected using integer or character indices, but with [] we can use an index vector to retrieve a slice with multiple members. 

For example:
```{r}
n=c(2,3,5)
s=c("aa","bb","cc","dd","ee")
b=c(TRUE,FALSE,TRUE,TRUE)
x=list(n,s,b,3)
x[[2]]
x[2]
```

    
    (b) Explain what the similarities and differences are between an `R
    data.frame` and an `R list`.
    
A: Data frames and List are similiar. The main difference is dataframe has more restrictions than list, the first is all elements of a dataframe are vectors,and another one is all elements of a dataframe have an equal length. This makes dataframe is always a two-dimensional structure which enables you to select rows and do operations on rows easier.
