---
title: "Trexquant-Mengyu Huang"
output: html_document
---

For R is more convenient with organizing data, I used R to program this exercise. However, the tradeoff is, the R program runs relatively slower than other languages.

I would like to simply state my algorithm first

1 read in all the dictionary words
2 choose one as the guessing word for the hangman game. And we would like to find out the optimal way of guessing the right answer
3 when we start playing the game, the player can only know about the length of the word we have to guess. So we could get a subset A using this specific word length. We start guessing by the top frequency letter in subset A
4 if our guess succeed, then we could get a new subset of words containing the letter we guessed
5 if our guess failed, then we could get a new subset of words without the letter we guessed
6 we start a new round of guessing by guess the top frequency letter in our new subset (Be careful that this new guessing letter should be different from what we have used. If the top frequency letter has been used, then we moved to use the 2nd highest frequency letter...)
7 keep repeating step 4,5,6, until the missing letters reach 6 or we have guessed out the right answer

The assumption for this algorithm is that letters with higher occurrence frequency has the larger probability to win. This algorithm is trying the most accurate way of winning the game. During the process, we need to update the subset all the way, so the tradeoff is that this algorithm runs a little slow. 

```{r}
#read the data into 'dictionary'. Split every word into separate letters and store the list into 'letter_table'. Count the number of letters of each word and store into 'letter_length'

dictionary<-as.matrix(read.table('words.txt'))
letter_table<-lapply(dictionary,strsplit,split="")
letter_length<-unlist(lapply(as.matrix(letter_table),lapply,length))

#we need to use the package 'stringr'. Install the package if haven't downloaded.
#install.packages('stringr')
library('stringr')
```

We need a helper function 'frequency' to help us generate the guess letters. This function is used to count the frequency of each letters in the input wordlist, and return the letter of the specific frequency order the user wanted. The default return value is the top frequency letter.


```{r}
frequency<-function(list,rank=1){
  letters=c('a','b','c','d','e','f','g','h','i','j','k','l','m','n','o','p','q','r','s','t','u','v','w','x','y','z')
  letters_count<-NULL
  
  for(i in 1:26){
    count=sum(sapply(list,function(x)sum(str_count(x,letters[i]))))
    letters_count<-c(letters_count,count)
  }
  
  letters_frequency=cbind.data.frame(letters,letters_count)
  pos<-order(letters_frequency[,2],decreasing=TRUE)[rank]
  fre<-as.character(letters_frequency[order(letters_frequency[,2],decreasing=TRUE)[rank],1])
  return(fre)
}
```

Write a function for the driver program. The inputs includes the word we want to guess, and the letter_table & letter_length we get before

```{r}

driver_program<-function(guess_word, letter_table, letter_length){
  
  #The first step of our algorithm would be creating a subset using the length of the guessword. When we start the hangman game, the only thing we know is the word length. So we create a subset A of the dictionary using the word-length, and start guessing from the top-frequency letter in A
  
  length<-nchar(guess_word)
  blanks<-rep('-',length)
  line<-c(blanks,'missed:','\n')
  cat(line)
  #create the subset
  subset_list<-letter_table[letter_length==length]
  subset_table<-dictionary[letter_length==length]
  guess_letter<-frequency(subset_list)
  guessed_letters=NULL
  missed_letters=NULL
  
  #after guessing the first letter, there are two conditions:
  #1. we've guessed the right letter. Under this circumstances, we could create the subset of words with the guessed letter.
  #2. we failed to guess the right letter. Under this circumstance, we could create the subset of words without the letter we guessed
  #3. The next guessing number would be the top frequency letter of our current subset. Also, we have to judge whether we have used this guessing letter before. If we had, then we choose the 2nd highest frequency letter...
  #4. keep repeating the process, until we have tried 6 times or we have guessed out the whole word.
  
  
  while(length(missed_letters)<6 && any(blanks!=unlist(strsplit(guess_word,"")))){
    line<-c('guess: ',guess_letter,'\n')
    cat(line)
    
    if(grepl(guess_letter,guess_word)){
      pos<-which(strsplit(guess_word,"")[[1]]==guess_letter)
      blanks[pos]=guess_letter
      subset_list<-subset_list[sapply(subset_table,grepl,pattern=guess_letter)]
      subset_table<-subset_table[sapply(subset_table,grepl,pattern=guess_letter)]
      line<-c(blanks,'missed: ', missed_letters,'\n')
    }else{
      subset_list<-subset_list[!sapply(subset_table,grepl,pattern=guess_letter)]
      subset_table<-subset_table[!sapply(subset_table,grepl,pattern=guess_letter)]
      missed_letters<-c(missed_letters,guess_letter)
      line<-c(blanks,'missed: ',missed_letters,'\n')
    }
    guessed_letters<-c(guessed_letters,guess_letter)
    cat(line)
    rank=1
    while(any(grepl(frequency(subset_list,rank),guessed_letters))){
      rank=rank+1
    }
    guess_letter<-frequency(subset_list,rank)
  }
  
  #we would like to know if we win the game or not
  if(all(unlist(strsplit(guess_word,""))==blanks)){
    win=TRUE
  }else{
    win=FALSE
  }
  return(win)
}

#using several word to test our driver program. I used 'abbatie', 'abbott' and 'alle' here
driver_program('abbatie',letter_table,letter_length)
driver_program('abbott',letter_table,letter_length)
driver_program('alle',letter_table,letter_length)
```
